Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2018-07-24 SupportingInformation.Rmd
Taylors law plot modified

#***constants
N<-100 #length of time series
n<-25 #number of sampling locations
numdat<-25 #the number of datasets to get from the normal copula
rho<-0.7 #normal-copula parameter for dependence between locations
          #this is a covariance, see below call to rmvnorm

#***generate the starting data, which has normal copula structure
#because it comes originally from a multivariate normal, before 
#transformation
sig<-matrix(rho,n,n)
diag(sig)<-1
m<-rmvnorm(N*numdat,mean=rep(0,n),sigma=sig)
m<-array(m,c(N,numdat,n))
m<-aperm(m,c(1,3,2))
gshape<-7.5
gscale<-1 #gamma parameters
cop_to_marg_trans<-function(x)
{
  return(qgamma(x,shape=gshape,scale=gscale)) #this is where we choose the marginals,
                                  #should be something skewed
}
marg_to_cop_trans<-function(x) #transforms to copula space
{
  return(pgamma(x,shape=gshape,scale=gscale))
}
m<-cop_to_marg_trans(pnorm(m))

#***surrogates using a Clayton copula, kendall preserving
#get the Clayton parameter to use
ncop<-normalCopula(rho,2)
ccop<-claytonCopula(2,2)
cparam<-iTau(ccop,tau(ncop))
ccop<-claytonCopula(cparam,n)

#generate the surrogates
csurrog<-array(NA,dim(m))
for (counter in 1:dim(m)[3])
{
  csurrog[,,counter]<-copsurrognd(m[,,counter],targetcop=ccop,numsurrog=1)
}

@@

#generate the surrogates
fsurrog<-array(NA,dim(m))
for (counter in 1:dim(m)[3])
{
  fsurrog[,,counter]<-copsurrognd(m[,,counter],targetcop=fcop,numsurrog=1)
}

@@

@@

#generate the surrogates
csurrog<-array(NA,dim(m))
for (counter in 1:dim(m)[3])
{
  csurrog[,,counter]<-copsurrognd(m[,,counter],targetcop=ccop,numsurrog=1)
}

@@

kend.stats.res<-list()
kend.points.res<-list()
spea.stats.res<-list()
spea.points.res<-list()
for (counter in 1:length(kend.surrog))
{
 ks<-kend.surrog[[counter]]
 ss<-spea.surrog[[counter]]
 kend.stats.res[[counter]]<-matrix(NA,5,numdat)
 spea.stats.res[[counter]]<-matrix(NA,5,numdat)
 kend.points.res[[counter]]<-array(NA,c(N,2,numdat))
 spea.points.res[[counter]]<-array(NA,c(N,2,numdat))
 for (sc in 1:numdat)
 {
   khold<-tl_stats(ks[,,sc])
   kend.stats.res[[counter]][,sc]<-khold$s.stats #spatial TL only
   rownames(kend.stats.res[[counter]])<-names(khold$s.stats)
   kend.points.res[[counter]][,,sc]<-as.matrix(khold$s.points)

@@

<!--Now make some plots, these one for kendall surrogates-->
```{r taylor_plots, echo=F}
##plot TL for clayton
xlimits<-range(kend.points.res[[1]][,1,],kend.points.res[[2]][,1,],kend.points.res[[3]][,1,],kend.points.res[[4]][,1,])
ylimits<-range(kend.points.res[[1]][,2,],kend.points.res[[2]][,2,],kend.points.res[[3]][,2,],kend.points.res[[4]][,2,])
par(mfrow=c(1,1),mar=c(4,4,1,1) + 0.1)
plot(kend.points.res[[1]][,1,],kend.points.res[[1]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")))  
points(kend.points.res[[1]][,1,1],kend.points.res[[1]][,2,1],type='p',pch=20,cex=.75,col='red')
points(kend.points.res[[1]][,1,2],kend.points.res[[1]][,2,2],type='p',pch=20,cex=.75,col='green')

##plot TL for normal copula
plot(kend.points.res[[2]][,1,],kend.points.res[[2]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")))  
points(kend.points.res[[2]][,1,1],kend.points.res[[2]][,2,1],type='p',pch=20,cex=.75,col='red')
points(kend.points.res[[2]][,1,2],kend.points.res[[2]][,2,2],type='p',pch=20,cex=.75,col='green')

#plot TL for frank
plot(kend.points.res[[3]][,1,],kend.points.res[[3]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")))  
points(kend.points.res[[3]][,1,1],kend.points.res[[3]][,2,1],type='p',pch=20,cex=.75,col='red')
points(kend.points.res[[3]][,1,2],kend.points.res[[3]][,2,2],type='p',pch=20,cex=.75,col='green')

#plot TL for survival clayton
plot(kend.points.res[[4]][,1,],kend.points.res[[4]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")))  
points(kend.points.res[[4]][,1,1],kend.points.res[[4]][,2,1],type='p',pch=20,cex=.75,col='red')
points(kend.points.res[[4]][,1,2],kend.points.res[[4]][,2,2],type='p',pch=20,cex=.75,col='green')

#prepare for plots coming below
s.lin<-cbind(kend.stats.res[[1]][1,],kend.stats.res[[2]][1,],kend.stats.res[[3]][1,],kend.stats.res[[4]][1,])
s.hom<-cbind(kend.stats.res[[1]][2,],kend.stats.res[[2]][2,],kend.stats.res[[3]][2,],kend.stats.res[[4]][2,])
s.rmse<-cbind(kend.stats.res[[1]][3,],kend.stats.res[[2]][3,],kend.stats.res[[3]][3,],kend.stats.res[[4]][3,])
s.int<-cbind(kend.stats.res[[1]][4,],kend.stats.res[[2]][4,],kend.stats.res[[3]][4,],kend.stats.res[[4]][4,])
s.slope<-cbind(kend.stats.res[[1]][5,],kend.stats.res[[2]][5,],kend.stats.res[[3]][5,],kend.stats.res[[4]][5,])

#plot s.lin distr for each set of surrogates
#***DAN: x axis needs to be redone to show the copula names used in the surrogates
par(mfrow=c(1,1),mar=c(4,4,1,1) + 0.1)
x<-1:4
plot(x,s.lin[1,],type="b",ylim=range(s.lin),xlab="Set of surrogates",ylab="Linearity test p-vals")
for (counter in 2:dim(s.lin)[1])
{
  lines(x,s.lin[counter,],type='b')
}

#plot s.hom distr for each set of surrogates
#***DAN: x axis needs to be redone to show the copula names used in the surrogates
x<-1:4
plot(x,s.hom[1,],type="b",ylim=range(s.hom),xlab="Set of surrogates",ylab="Homoskedasticity test p-vals")
for (counter in 2:dim(s.slope)[1])
{
  lines(x,s.hom[counter,],type='b')
}

#plot s.rmse distr for each set of surrogates
#***DAN: x axis needs to be redone to show the copula names used in the surrogates
x<-1:4
plot(x,s.rmse[1,],type="b",ylim=range(s.rmse),xlab="Set of surrogates",ylab="Spatial TL rmse")
for (counter in 2:dim(s.rmse)[1])
{
  lines(x,s.rmse[counter,],type='b')
}

#plot s.int distr for each set of surrogates
#***DAN: x axis needs to be redone to show the copula names used in the surrogates
x<-1:4
plot(x,s.int[1,],type="b",ylim=range(s.int),xlab="Set of surrogates",ylab="Spatial TL intercept")
for (counter in 2:dim(s.int)[1])
{
  lines(x,s.int[counter,],type='b')
}

#plot s.slope distr for each set of surrogates
#***DAN: x axis needs to be redone to show the copula names used in the surrogates
x<-1:4
plot(x,s.slope[1,],type="b",ylim=range(s.slope),xlab="Set of surrogates",ylab="Spatial TL slope")
for (counter in 2:dim(s.slope)[1])
{
  lines(x,s.slope[counter,],type='b')
}
```

>>>>

#***constants
N<-50 #length of time series
n<-25 #number of sampling locations
numdat<-10 #the number of datasets to get from the normal copula
rho<-0.7 #normal-copula parameter for dependence between locations
          #this is a covariance, see below call to rmvnorm

#***generate the starting data, which has normal copula structure
#because it comes originally from a multivariate normal, before 
#transformation
sig<-matrix(rho,n,n)
diag(sig)<-1
m<-rmvnorm(N*numdat,mean=rep(0,n),sigma=sig) # each column of m has normal histogram
m<-array(m,c(N,numdat,n))
m<-aperm(m,c(1,3,2)) # 2nd dim is normally distributed e.g. hist(m[,1,])
gshape<-7.5
gscale<-1 #gamma parameters

cop_to_marg_trans<-function(x){
  return(qgamma(x,shape=gshape,scale=gscale)) #this is where we choose the marginals,
                                  #should be something skewed
}

marg_to_cop_trans<-function(x){ #transforms to copula space
  return(pgamma(x,shape=gshape,scale=gscale))
}

# pnorm(m) makes normal to uniform
# then qgamma(...) makes uniform to gamma distributed
m<-cop_to_marg_trans(pnorm(m))

#***surrogates using a Clayton copula, kendall preserving

#get the Clayton parameter to use
ncop<-normalCopula(rho,2)
ccop<-claytonCopula(2,2)
cparam<-iTau(ccop,tau(ncop))
ccop<-claytonCopula(cparam,n)

#generate the surrogates
csurrog<-array(NA,dim(m))
for (counter in 1:dim(m)[3]){
  csurrog[,,counter]<-copsurrognd(m[,,counter],targetcop=ccop,numsurrog=1)
}

@@

#generate the surrogates
fsurrog<-array(NA,dim(m))
for (counter in 1:dim(m)[3]){
  fsurrog[,,counter]<-copsurrognd(m[,,counter],targetcop=fcop,numsurrog=1)
}

@@

# check : plot(csurrog[,1,1],csurrog[,2,1]) and plot(scsurrog[,1,1],scsurrog[,2,1])

@@

#generate the surrogates
csurrog<-array(NA,dim(m))
for (counter in 1:dim(m)[3]){
  csurrog[,,counter]<-copsurrognd(m[,,counter],targetcop=ccop,numsurrog=1)
}

@@

kend.stats.res<-list()
kend.points.res<-list()
spea.stats.res<-list()
spea.points.res<-list()
for (counter in 1:length(kend.surrog)){
 ks<-kend.surrog[[counter]]
 ss<-spea.surrog[[counter]]
 kend.stats.res[[counter]]<-matrix(NA,5,numdat) # to keep spatial TL stats
 spea.stats.res[[counter]]<-matrix(NA,5,numdat)
 kend.points.res[[counter]]<-array(NA,c(N,2,numdat)) # to keep log(mu), log(var) of spatial TL as a dataframe
 spea.points.res[[counter]]<-array(NA,c(N,2,numdat))
 for (sc in 1:numdat){
   khold<-tl_stats(ks[,,sc])
   kend.stats.res[[counter]][,sc]<-khold$s.stats #spatial TL only
   rownames(kend.stats.res[[counter]])<-names(khold$s.stats)
   kend.points.res[[counter]][,,sc]<-as.matrix(khold$s.points)

@@

<!--Now make some plots, these one for kendall surrogates-->
```{r taylor_plots, echo=F, results="hide"}
xlimits<-range(kend.points.res[[1]][,1,],kend.points.res[[2]][,1,],kend.points.res[[3]][,1,],kend.points.res[[4]][,1,])
ylimits<-range(kend.points.res[[1]][,2,],kend.points.res[[2]][,2,],kend.points.res[[3]][,2,],kend.points.res[[4]][,2,])

pdf("./Results/taylorslaw_results/SpatialTL_CNFSC.pdf",height=4,width=16)
op<-op<-par(mfrow=c(1,4),mar=c(4,5,1,1),mgp=c(3,1,0))
##plot TL for clayton
plot(kend.points.res[[1]][,1,],kend.points.res[[1]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
     cex.axis=1.5,cex.lab=1.5)  
points(kend.points.res[[1]][,1,1],kend.points.res[[1]][,2,1],type='p',pch=20,cex=.75,col='red')
points(kend.points.res[[1]][,1,2],kend.points.res[[1]][,2,2],type='p',pch=20,cex=.75,col='green')
mtext(paste("Clayton"),side = 3, line=-2, adj=0.1, col="black")

##plot TL for normal copula
plot(kend.points.res[[2]][,1,],kend.points.res[[2]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
     cex.axis=1.5,cex.lab=1.5)  
points(kend.points.res[[2]][,1,1],kend.points.res[[2]][,2,1],type='p',pch=20,cex=.75,col='red')
points(kend.points.res[[2]][,1,2],kend.points.res[[2]][,2,2],type='p',pch=20,cex=.75,col='green')
mtext(paste("Normal"),side = 3, line=-2, adj=0.1, col="black")

#plot TL for frank
plot(kend.points.res[[3]][,1,],kend.points.res[[3]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
     cex.axis=1.5,cex.lab=1.5)  
points(kend.points.res[[3]][,1,1],kend.points.res[[3]][,2,1],type='p',pch=20,cex=.75,col='red')
points(kend.points.res[[3]][,1,2],kend.points.res[[3]][,2,2],type='p',pch=20,cex=.75,col='green')
mtext(paste("Frank"),side = 3, line=-2, adj=0.1, col="black")

#plot TL for survival clayton
plot(kend.points.res[[4]][,1,],kend.points.res[[4]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
     cex.axis=1.5,cex.lab=1.5)  
points(kend.points.res[[4]][,1,1],kend.points.res[[4]][,2,1],type='p',pch=20,cex=.75,col='red')
points(kend.points.res[[4]][,1,2],kend.points.res[[4]][,2,2],type='p',pch=20,cex=.75,col='green')
mtext(paste("Survival Clayton"),side=3, line=-2, adj=0.1, col="black")

par(op)
dev.off()
#---------------------------------------------------
#prepare for plots coming below
s.lin<-cbind(kend.stats.res[[1]][1,],kend.stats.res[[2]][1,],kend.stats.res[[3]][1,],kend.stats.res[[4]][1,])
s.hom<-cbind(kend.stats.res[[1]][2,],kend.stats.res[[2]][2,],kend.stats.res[[3]][2,],kend.stats.res[[4]][2,])
s.rmse<-cbind(kend.stats.res[[1]][3,],kend.stats.res[[2]][3,],kend.stats.res[[3]][3,],kend.stats.res[[4]][3,])
s.int<-cbind(kend.stats.res[[1]][4,],kend.stats.res[[2]][4,],kend.stats.res[[3]][4,],kend.stats.res[[4]][4,])
s.slope<-cbind(kend.stats.res[[1]][5,],kend.stats.res[[2]][5,],kend.stats.res[[3]][5,],kend.stats.res[[4]][5,])

pdf("./Results/taylorslaw_results/SpatialTL_5stats.pdf",height=3,width=15)
op<-par(mfrow=c(1,5),mar=c(4.2,5,1,1),mgp=c(3,1,0))
#***DAN: x axis needs to be redone to show the copula names used in the surrogates

x<-1:4
x2<-c("C", "N", "F", "SC")
plot(x,s.lin[1,],type="l",col="grey",ylim=range(s.lin),xlab="Set of surrogates",ylab="Linearity test p-vals",
     cex.axis=1.5,cex.lab=1.5,xaxt="n")
points(x,s.lin[1,],col="red")
axis(1, at = c(1:4), x2, cex.axis=1.5)
for (counter in 2:dim(s.lin)[1]){
  lines(x,s.lin[counter,],col="grey")
  points(x,s.lin[counter,],col="red")
}

#plot s.hom distr for each set of surrogates
#***DAN: x axis needs to be redone to show the copula names used in the surrogates
x<-1:4
plot(x,s.hom[1,],type="l",col="grey",ylim=range(s.hom),xlab="Set of surrogates",
     ylab="Homoskedasticity test p-vals",
     cex.axis=1.5,cex.lab=1.5,xaxt="n")
points(x,s.hom[1,],col="red")
axis(1, at = c(1:4), x2, cex.axis=1.5)
for (counter in 2:dim(s.slope)[1]){
  lines(x,s.hom[counter,],col="grey")
  points(x,s.hom[counter,],col="red")
}

#plot s.rmse distr for each set of surrogates
x<-1:4
plot(x,s.rmse[1,],type="l",col="grey",ylim=range(s.rmse),xlab="Set of surrogates",ylab="Spatial TL rmse",
     cex.axis=1.5,cex.lab=1.5,xaxt="n")
points(x,s.rmse[1,],col="red")
axis(1, at = c(1:4), x2, cex.axis=1.5)
for (counter in 2:dim(s.rmse)[1]){
  lines(x,s.rmse[counter,],col="grey")
  points(x,s.rmse[counter,],col="red")
}

#plot s.int distr for each set of surrogates
#***DAN: x axis needs to be redone to show the copula names used in the surrogates
x<-1:4
plot(x,s.int[1,],type="l",col="grey",ylim=range(s.int),xlab="Set of surrogates",
     ylab="Spatial TL intercept",
     cex.axis=1.5,cex.lab=1.5,xaxt="n")
points(x,s.int[counter,],col="red")
axis(1, at = c(1:4), x2, cex.axis=1.5)
for (counter in 2:dim(s.int)[1]){
  lines(x,s.int[counter,],col="grey")
  points(x,s.int[counter,],col="red")
}

#plot s.slope distr for each set of surrogates
#***DAN: x axis needs to be redone to show the copula names used in the surrogates
x<-1:4
plot(x,s.slope[1,],type="l",col="grey",ylim=range(s.slope),xlab="Set of surrogates",
     ylab="Spatial TL slope",
     cex.axis=1.5,cex.lab=1.5,xaxt="n")
points(x,s.slope[counter,],col="red")
axis(1, at = c(1:4), x2, cex.axis=1.5)
for (counter in 2:dim(s.slope)[1]){
  lines(x,s.slope[counter,],col="grey")
  points(x,s.slope[counter,],col="red")
}
par(op)
dev.off()
```

\begin{figure}[!h]
\textbf{ \hspace{3.4 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.1 cm} (D)} \\
\vspace{-0.5 cm} 
\begin{center}
\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_CNFSC.pdf}\\
\end{center}
\textbf{ \hspace{3.1 cm} (E) \hspace{2.4 cm} (F) \hspace{2.3 cm} (G) \hspace{2.3 cm} (H) \hspace{2.3 cm} (I)}\\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_5stats.pdf}
\caption[Spatial Taylor's Law results]{Spatial Taylor's Law results (A) hjghj ; (B) fgjfgjfg. \label{fig_spTL}}
\end{center}
\end{figure}
